### 2015-05-27 doc/manual/mscviewer.tex
major changes to build infra and streamlining of python API
fixed user guide to just mention python-api guide
V2.0.0

%\usepackage[margin=1in]{geometry}
\usepackage{fixltx2e} 
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage{centernot}
\usepackage{listings}
\usetikzlibrary{shadows}
\usepackage[toc,page]{appendix}
\usepackage{etoolbox}
\usepackage{wrapfig}
\usepackage{tcolorbox}
\usepackage{lipsum}
\include{titlepage}

@@

\newcommand{\todo}[1]{\textit{#1}}
\newcommand{\defterm}[1]{\textit{#1}\index{#1}}

%\title{\prog{} v1.0} \author{ Roberto Attias  \\
%    User Manual
%    }
\date{\today}

@@

\lstset{frame=tb,
  language=Python,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=left,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{mauve},
  stringstyle=\color{dkgreen},
  backgroundcolor=\color{lightcyan},
  frame=shadowbox,
  xleftmargin=2em,
  framexleftmargin=1.5em, 
  rulesepcolor=\color{gray},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3,
  escapeinside={(*}{*)}
}

@@

\section{\prog}
\prog{} is a tool supporting visualization and analysis of message sequence
charts. The tool loads a text file, interprets lines containing the
\mscevent{} or \mscentity{} keywords and representing events and interactions occurring on
entities, ignoring other lines. From the interpreted lines a model is
constructed in memory. In the GUI the tool can visualize the list of enties,
from which the user can select a subset to be shown in a sequence diagram chart.
through keyboard or mouse commands the user can easily navigate through events
and interactions, search, and perform a number of other actions for browsing the
diagram.

In a design phase the user can write a concise description of one or
more flows in the inpujt language,load it in the tool, and export an image to 
be used for documentation.

By instrumenting certain point in the program or application with syslogs,
printf, traces or any other mechanism avaliable, the user can generate run the
application and generate the input file for \prog{}. In this case the input
file contains flows corresponding to actual executions, which could be visually
inspected for verification.

Visual inspection of flows from execution logs is useful, but doesn't scale for
debugging, as it requires a human operator familiar with the expected flows.
When applications are large enough, their development might be distributed
across multiple teams, with only few members fully aware of cross-component
flows. To address this problem \prog{} is integrated with a Python interpreter
and supports execution of scripts which can browse the model and perform any
sort of validation. In addition, a set of classes is provided to support a
concise description of expected flows. With these mechanism the user can create
validation script to be executed in the tool on the model created from execution
logs, automatically validating their correctness. Scripts can also be executed
from command line, without any GUI visualization, supporting batch validations
and product smoke-testing.

@@

When designing, implementing and debugging such a complex system flows play a
very important role in the entire development cycle.

@@

It's worth mentioning that \prog{} does not have any semantic knowledge of
concept such as nodes, processes, or threads, and the model could be used to
represent any kind of problem domain. \prog{} understands only the concept of
hyerarchical IDS and names where elements in the hyerarchy are separated by
slashes.

@@

>>>>

%\usepackage[margin=1in]{geometry}
\usepackage{fixltx2e} 
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage{centernot}
\usepackage{listings}
\usetikzlibrary{shadows}
\usepackage[toc,page]{appendix}
\usepackage{etoolbox}
\usepackage{wrapfig}
\usepackage{tcolorbox}
\usepackage{lipsum}
\include{generated/titlepage}

@@

\newcommand{\todo}[1]{\textit{#1}}
\newcommand{\defterm}[1]{\textit{#1}\index{#1}}
%
\makeatletter
\newenvironment{indentation}[2]%
{ \par \setlength{\leftmargin}{#1}
  \setlength{\rightmargin}{#2}
  \advance\linewidth -\leftmargin
  \advance\linewidth -\rightmargin
  \advance\@totalleftmargin\leftmargin
  \@setpar{{\@@par}}%
  \parshape 1 \@totalleftmargin
  \linewidth \ignorespaces
}{\par}
\makeatother
%
%\title{\prog{} v1.0} \author{ Roberto Attias  \\
%    User Manual
%    }
\date{\today}

@@

\lstset{frame=tb,
  language=Python,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=left,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{mauve},
  stringstyle=\color{dkgreen},
  %backgroundcolor=\color{lightcyan},
  %frame=shadowbox,
  %xleftmargin=2em,
  %framexleftmargin=1.5em, 
  %rulesepcolor=\color{gray},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3,
  escapeinside={(*}{*)}
}

@@

\section{\prog}
\prog{} is a tool supporting visualization and analysis of message sequence
charts intended primaritly for debugging of logs/traces obtained from concurrent systems.

The tool loads a text file and interprets lines containing well known keywords representing
events and interactions occurring in entities. Lines not containing the known keywords are
ignored. From the interpreted lines a model is constructed in memory. In the GUI the tool 
shows the list of all entities for which events exist. From this list the user can 
select a subset to be shown in a sequence diagram chart. Through keyboard or mouse commands
the user can easily navigate through events and interactions, rearrange entities, search, 
and perform a number of other actions for browsing the diagram.

In addition to using the tool for visualization of traces for existing 
systems, during a design phase the user can write a concise description of one or
more flow\footnote{a formal definiton of flow will be provided later in this document.
For the time being consider it a sequence of events across one or more entities} 
in the input language, load it in the tool, and export an image to 
be used for documentation. Later on, the implemented system should produce diagrams
similar to the one defined at design time, hence this approach can be used to confirm
the system behaves as designed.

In order for a system to produce the required data for \prog{}, the program code 
need to be instrumented with syslogs, printf, traces or any other mechanism avaliable,
using the format expected by the tool for the messages.
The user can run the application and generate the input file for \prog{}, possibly
by merging together multiple generated files.

Visual inspection of the system behavior from execution logs is useful, but doesn't
scale for debugging, as it requires a human operator familiar with the expected behavior.
When applications are large enough, their development might be distributed
across multiple teams, and possibly even geographically, with only few members
fully aware of cross-component flows. To address this problem \prog{} is 
integrated with a Python interpreter and supports execution of scripts which 
can browse the model and perform any sort of validation. In addition, a set
of classes is provided to support a concise description of expected flows.
With these mechanism the user can create script to be executed in
the tool on the model created from execution logs, automatically validating 
their correctness. Scripts can also be executed from command line, without 
any GUI visualization, supporting batch validations and product smoke-testing.

@@

When designing, implementing and debugging such a complex system, flows play a
very important role in the entire development cycle.

@@

It's worth mentioning that \prog{} does not have any semantic knowledge of
concept such as nodes, processes, or threads, and the model could be used to
represent any kind of problem domain. \prog{} understands only the concept of
hierarchical IDS and names where elements in the hierarchy are separated by
slashes.

@@

\chapter{Pyhton API}
One of the most powerful aspects of \prog{} is the integration with a Python interpreter
which allows to access programmatically both model as wel as GUI.

One use of this feature
consist in writing validation scripts that can browse model data and report whether a 
particular computational flow is present or not, it respects certain timing constraints, etc. 
Because the script is written once and then applied to traces produced by running the 
system multiple times, this scales the effort of an activity that otherwise would have to be
performed by one or more human beings.

Python scripts can also be used to drive the GUI. For example, the screen captures in this 
manual are all generated automatically a build time by a Python script which loads a model,
open some entities in the sequence diagram, captures screenshots and saves them.

The interaction between the user Python script and the tool is done through APIs 
collected in a few Python files the script can import. The Python language doesn't 
have a concept of private vs. public API, and the technology used to integrate the 
Python interpreted in \prog{} (which is written in Java) results in all the Java 
classes being actually visible and accessible from the Python code. However the 
user should restrict itself to using the API documented in the next few sections,
as this provides better guarantees of not being modified in the future. No such 
claims are made agains any of the Java classes.

The API is currently divided in the following modules:
\begin{itemize}
\item \texttt{msc.model} -- provides API to access entities, events and interactions in the model
\item \texttt{msc.gui} -- provides API to access the GUI 
\item \texttt{msc.utils} -- provides some utility API 
\item \texttt{msc.graph} -- provides API to access the graphing capabilities (experimental). 
\item \texttt{msc.flowdef} -- provides API to implement compact definition of expected flows and apply them
                              to the current model
\end{itemize}                              
\section{Model API}

\input{generated/python-api-model.tex}

\section{GUI API}
\input{generated/python-api-gui.tex}

\section{Flows API}
\input{generated/python-api-flowdef.tex}

\section{Graph API}
\input{generated/python-api-graph.tex}
